---
title: "Homework 2"
output: github_document
date: "2023-09-28"
---

```{r}
library(tidyverse)
library(readxl)
```


## Problem 1

clean the data in pols-month.csv

```{r}
pols_df = 
  read_csv("data/pols-month.csv") |>
  janitor::clean_names() |>
  separate(mon, into = c("year", "month","day")) |>
  mutate(
    month = case_match(
      month,
      "01" ~ "Jan",
      "02" ~ "Feb",
      "03" ~ "Mar",
      "04" ~ "Apr",
      "05" ~ "May",
      "06" ~ "Jun",
      "07" ~ "Jul",
      "08" ~ "Aug",
      "09" ~ "Sep",
      "10" ~ "Oct",
      "11" ~ "Nov",
      "12" ~ "Dec"
    ),
    year = as.numeric(year),
    prez_dem = replace(prez_dem,prez_dem == "1", "dem"),
    president = replace(prez_dem,prez_dem == "0", "gop")
    ) |>
  select( -prez_dem, -prez_gop, -day)

```

clean the data in snp.csv

```{r}
snp_df = 
  read_csv("data/snp.csv") |>
  janitor::clean_names() |>
  separate(date, into = c("month", "day","year")) |>
  mutate(
    year = as.numeric(year),
    year = ifelse( year <= 15, year + 2000, year + 1900),
    month = as.numeric(month)
  ) |>
  arrange(year, month) |>
  mutate(
    month = as.character(month),
    month = case_match(
      month,
      "1" ~ "Jan",
      "2" ~ "Feb",
      "3" ~ "Mar",
      "4" ~ "Apr",
      "5" ~ "May",
      "6" ~ "Jun",
      "7" ~ "Jul",
      "8" ~ "Aug",
      "9" ~ "Sep",
      "10" ~ "Oct",
      "11" ~ "Nov",
      "12" ~ "Dec"
    ),
  ) |>
  select(year, month, close) 

```

tidy the unemployment data

```{r}
unemployment_df = 
  read_csv("data/unemployment.csv") |>
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "percentage") |>
  janitor::clean_names()

```

join the datasets

```{r}
combine_df = 
  full_join(snp_df,pols_df)
combine_df = 
  full_join(combine_df, unemployment_df)
```

The dataset `pols_df` has `r nrow(pols_df)` observations and `r ncol(pols_df)` variables. The range of years is from `r pols_df$year[1]` to `r pols_df$year[nrow(pols_df)]`. The data shows the president's political side andthe number of national politicians who are either democratic or republican at the time stated.

The dataset `snp_df` has `r nrow(snp_df)` observations and `r nrow(snp_df)` variables. The range of years is from `r snp_df$year[1]` to `r snp_df$year[nrow(snp_df)]`. The data shows the closing values of the S&P stock index on the associated date.

The dataset `unemployment_df` has `r nrow(unemployment_df)` observations and `r nrow(unemployment_df)` variables. The range of years is from `r snp_df$year[1]` to `r unemployment_df$year[nrow(unemployment_df)]`. The data shows the percentage of unemployment on the associated date.


## Problem 2

clean Mr.Trash Wheel sheet

```{r}
trash_wheel_df =
  read_excel ("data/202207 Trash Wheel Collection Data.xlsx", 
              sheet = "Mr. Trash Wheel") |>
  select(-...15, -...16) |>
  janitor::clean_names() |>
  head(-1) |>
  mutate(
    homes_powered = 
      weight_tons * 500 / 30,
    Trash_Wheel = "Mr. Trash Wheel",
    year = as.numeric(year)
  )
```

clean Professor Trash Wheel sheet

```{r}
professor_trash_df =
  read_excel ("data/202207 Trash Wheel Collection Data.xlsx", 
              sheet = "Professor Trash Wheel") |>
  janitor::clean_names() |>
  head(-1) |>
  mutate(
    homes_powered = 
      weight_tons * 500 / 30,
    Trash_Wheel = "Professor Trash Wheel"
  )
```

clean Gwynnda Trash Wheel sheet

```{r}
gwynnda_trash_df =
  read_excel ("data/202207 Trash Wheel Collection Data.xlsx", 
              sheet = "Gwynnda Trash Wheel") |>
  janitor::clean_names() |>
  head(-2) |>
  mutate(
    homes_powered = 
      weight_tons * 500 / 30,
    Trash_Wheel = "Gwynnda Trash Wheel"
  )
```

combine sheets

```{r}
trash_wheel_collection = 
  bind_rows(trash_wheel_df, professor_trash_df, gwynnda_trash_df)
```

The dataset `trash_wheel_df` has `r nrow(trash_wheel_df)` observations and `r ncol(trash_wheel_df)` variables. The `weight_tons` is the weight of trash in tons. The `homes_powered` is calculated using trash `weight_tons`. On `r trash_wheel_df$date[1]`, dumpster `r trash_wheel_df$dumpster[1]` has `r trash_wheel_df$weight_tons[1]` tons of trash, which could power average `r trash_wheel_df$homes_powered[1]` homes.

The dataset `professor_trash_df` has `r nrow(professor_trash_df)` observations and `r ncol(professor_trash_df)` variables. The total weight of trash collect by Professor Trash Wheel is `r sum(professor_trash_df$weight_tons)` tons.

The dataset `gwynnda_trash_df` has `r nrow(gwynnda_trash_df)` observations and `r ncol(gwynnda_trash_df)` variables. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(filter(gwynnda_trash_df, month == "July" & year == 2021)$cigarette_butts)`.

The combined dataset `trash_wheel_collection` has `r nrow(trash_wheel_collection)` observations and `r ncol(trash_wheel_collection)` variables.












